Effectiveness of BNT162b2 booster doses in England: a observational study in OpenSAFELY-TPP: Materials

### 2022-05-04 manuscript/reports.Rmd
manuscript updates

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmitted",
      #"covidcc",
      "coviddeath",
      "noncoviddeath"
    ) %>% 
    set_names(., .)

@@

meta2hazards <- 
  # derive directly from hazards as meta2effects table needs to be re-run
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
    read_csv(fs::path(output_dir_os, "cev", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "age65plus", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "prior_covid_infection", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
  ) %>%
  filter(
    outcome %in% outcomes
  ) %>%
  mutate(
    period = (term_right>28) + 1,
    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
  ) %>%
  group_by(
    treatment, treatment_descr, outcome, outcome_descr, outcome_descr_wrap, subgroup, subgroup_variable, subgroup_descr, subgroup_level, subgroup_level_descr, model, model_descr, period, 
  ) %>%
  summarise(
    term_left = min(term_left),
    term_right = max(term_right),
    term_midpoint = (term_left+term_right)/2,
    estimate = weighted.mean(estimate, robust.se^-2),
    std.error = sqrt(1/sum(std.error^-2)),
    robust.se = sqrt(1/sum(robust.se^-2)),
    statistic = estimate/robust.se,
    p.value = 2 * pmin(pnorm(statistic), pnorm(-statistic)),
    conf.low = estimate + qnorm(0.025)*robust.se,
    conf.high = estimate + qnorm(0.975)*robust.se,
    hr = exp(estimate),
    hr.ll = exp(conf.low),
    hr.ul = exp(conf.high),
    ve = 1-hr,
    ve.ll = 1-hr.ul,
    ve.ul = 1-hr.ll
  ) %>%
  ungroup() %>%
  mutate(
    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
  )

# meta2hazards <- 
#   bind_rows(
#     read_csv(fs::path(output_dir_os, "none", "meta2effects.csv")),
#     read_csv(fs::path(output_dir_os, "vax12_type", "meta2effects.csv")),
#     read_csv(fs::path(output_dir_os, "cev", "meta2effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
#     read_csv(fs::path(output_dir_os, "age65plus", "meta2effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
#   ) %>%
#   filter(
#     outcome %in% outcomes
#   ) %>%
#   mutate(
#     subgroup_descr = fct_inorder(subgroup_descr),
#     subgroup_level = fct_inorder(subgroup_level),
#     subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
#     treatment_descr = fct_inorder(treatment_descr),
#     outcome_descr = fct_inorder(outcome_descr),
#     outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
#     #model_name = fct_inorder(str_wrap(model_name, 15)),
#     hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
#     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
#   )

hazards3 <- hazards %>% filter(model==3)

@@

### Figure 2: Estimated booster effectiveness (5 period)

@@

<!-- #### Figure 2b: Estimated booster effectiveness (2 period split at 28 days) -->

<!-- ```{r plot_effects28, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'} -->

<!--   postbaselinecuts28 <- c(0,28,70) -->

<!--   plot_hazards <- -->
<!--     meta28hazards %>% -->
<!--     filter( -->
<!--       model==3, -->
<!--       outcome %in% outcomes, -->
<!--       treatment %in% c("pfizer"), -->
<!--       subgroup_variable %in% c("none", "vax12_type") -->
<!--     ) %>% -->
<!--     ggplot()+ -->
<!--     geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
<!--     geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
<!--     geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts28+14)), colour="darkgrey") + -->
<!--     geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts28+14)), colour="darkgrey") + -->
<!--     geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts28+14)-7, xmax=max(postbaselinecuts28+14)+7, fill="grey95", alpha=0.5, colour="transparent") + -->
<!--     geom_hline(aes(yintercept=1), colour='grey')+ -->
<!--     facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+ -->
<!--     scale_y_log10( -->
<!--       breaks=y_breaks, -->
<!--       limits = c(0.01, 1.5), -->
<!--       oob = scales::oob_keep, -->
<!--       sec.axis = sec_axis( -->
<!--         ~(1-.), -->
<!--         name="Effectiveness (%)", -->
<!--         breaks = 1-(y_breaks), -->
<!--         labels = function(x){formatpercent100(x, 1)} -->
<!--       ) -->
<!--     )+ -->
<!--     scale_x_continuous(breaks=c(postbaselinecuts28, max(postbaselinecuts28+14)), labels=c(postbaselinecuts28, "Overall"), limits=c(min(postbaselinecuts28), max(postbaselinecuts28+14)+14), expand = c(0, 0))+ -->
<!--     scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+ -->
<!--     labs( -->
<!--       y="Hazard ratio", -->
<!--       x="Days since booster dose", -->
<!--       colour=NULL -->
<!--      ) + -->
<!--     theme_minimal()+ -->
<!--     theme( -->
<!--       panel.border = element_blank(), -->
<!--       axis.line.y = element_line(colour = "black"), -->
<!--       axis.line.y.right = element_blank(), -->

<!--       legend.position = "none" -->
<!--      ) + -->
<!--     NULL -->
<!--   plot_hazards -->

>>>>

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmittedproxy1",
      "covidadmitted",
      #"covidcc",
      "coviddeath",
      "noncoviddeath"
    ) %>% 
    set_names(., .)

@@

meta28hazards <- 
  # derive directly from hazards as meta2effects table needs to be re-run
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
    read_csv(fs::path(output_dir_os, "cev", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "age65plus", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "prior_covid_infection", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
  ) %>%
  filter(
    outcome %in% outcomes
  ) %>%
  mutate(
    period = (term_right>28) + 1,
    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
  ) %>%
  group_by(
    treatment, treatment_descr, outcome, outcome_descr, outcome_descr_wrap, subgroup, subgroup_variable, subgroup_descr, subgroup_level, subgroup_level_descr, model, model_descr, period, 
  ) %>%
  summarise(
    term_left = min(term_left),
    term_right = max(term_right),
    term_midpoint = (term_left+term_right)/2,
    estimate = weighted.mean(estimate, robust.se^-2),
    std.error = sqrt(1/sum(std.error^-2)),
    robust.se = sqrt(1/sum(robust.se^-2)),
    statistic = estimate/robust.se,
    p.value = 2 * pmin(pnorm(statistic), pnorm(-statistic)),
    conf.low = estimate + qnorm(0.025)*robust.se,
    conf.high = estimate + qnorm(0.975)*robust.se,
    hr = exp(estimate),
    hr.ll = exp(conf.low),
    hr.ul = exp(conf.high),
    ve = 1-hr,
    ve.ll = 1-hr.ul,
    ve.ul = 1-hr.ll
  ) %>%
  ungroup() %>%
  mutate(
    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
  )
# 
# meta28hazards <-
#   bind_rows(
#     read_csv(fs::path(output_dir_os, "none", "meta2effects.csv")),
#     read_csv(fs::path(output_dir_os, "vax12_type", "meta2effects.csv")),
#     read_csv(fs::path(output_dir_os, "cev", "meta2effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
#     read_csv(fs::path(output_dir_os, "age65plus", "meta2effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
#   ) %>%
#   filter(
#     outcome %in% outcomes
#   ) %>%
#   mutate(
#     subgroup_descr = fct_inorder(subgroup_descr),
#     subgroup_level = fct_inorder(subgroup_level),
#     subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
#     treatment_descr = fct_inorder(treatment_descr),
#     outcome_descr = fct_inorder(outcome_descr),
#     outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
#     #model_name = fct_inorder(str_wrap(model_name, 15)),
#     hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
#     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
#   )
# 
# hazards3 <- hazards %>% filter(model==3)

@@

### Figure 2a: Estimated booster effectiveness (5 period)

@@

#### Figure 2b: Estimated booster effectiveness (2 period split at 28 days)

```{r plot_effects28, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

  if_else(
    formatx==scales::label_percent(accuracy, suffix="")(1),
    paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)),
    formatx
  )
}
 ## hazard ratio

  overallhazards_plotdata <-
    overallhazards %>%
    filter(
      model==3,
      outcome %in% outcomes,
      #treatment %in% c("pfizer"),
      subgroup_variable %in% c("none", "vax12_type", "cev", "age65plus", "prior_covid_infection")
    ) %>% mutate(
      subgroup_level_descr_wrap=fct_relabel(subgroup_level_descr, ~str_wrap(.x, width=10))
    )

  postbaselinecuts28 <- c(0,28,70)

  plot_hazards <-
    meta28hazards %>%
    filter(
      model==3,
      outcome %in% outcomes,
      #treatment %in% c("pfizer"),
      subgroup_variable %in% c("none", "vax12_type", "cev", "age65plus", "prior_covid_infection")
    ) %>%
    mutate(
      subgroup_level_descr_wrap=fct_relabel(subgroup_level_descr, ~str_wrap(.x, width=10))
    ) %>%
    ggplot()+
    geom_point(aes(y=hr, x=term_midpoint, colour=treatment_descr), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=treatment_descr), position = position_dodge(width = 1.8))+
    geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts28+14), colour=treatment_descr), position = position_dodge(width = 1.8)) +
    geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts28+14), colour=treatment_descr), position = position_dodge(width = 1.8)) +
    geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts28+14)-7, xmax=max(postbaselinecuts28+14)+7, fill="grey95", alpha=0.4, colour="transparent") +
    geom_hline(aes(yintercept=1), colour='grey')+
    facet_grid(rows=vars(subgroup_level_descr_wrap), cols=vars(outcome_descr_wrap), switch="y")+
    scale_y_log10(
      breaks=y_breaks,
      limits = c(0.05, 1.5),
      oob = scales::oob_keep,
      sec.axis = sec_axis(
        ~(1-.),
        name="Effectiveness (%)",
        breaks = 1-(y_breaks),
        labels = function(x){formatpercent100(x, 1)}
      )
    )+
    scale_x_continuous(breaks=c(postbaselinecuts28, max(postbaselinecuts28+14)), labels=c(postbaselinecuts28, ""), limits=c(min(postbaselinecuts28), max(postbaselinecuts28+14)+14), expand = c(0, 0))+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
    labs(
      y="Hazard ratio",
      x="Days since booster dose",
      colour=NULL
     ) +
    theme_minimal()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.y.right = element_blank(),

      legend.position = "bottom"
     ) +
    NULL
  plot_hazards
